Parameter Space Representation of Pareto Front to Explore Hardware-Software Dependencies

### 2014-07-27 algorithm.tex
AA: aggiustato fino a initialization phase

\section{The PS algorithm}
\label{sec:algorithm}
In this section we present PS, the algorithm that we propose to perform the Design Space Exploration (DSE). As any other DSE algorithm, its input is the parameterized system $S$ and a benchmark application $b$, while the output is an approximation of the Pareto-optimal set $\mathcal{P}\left(S,b\right)$.   
PS is iterative and we will describe, separately, the initialization phase, the operations performed inside each iteration and the condition that triggers the termination.

@@

Each iteration is called \emph{era}. The $i$-th era is characterized by: 
	\begin{itemize}
	\item the set of regions $\mathcal{R}_{i}$ in which the configuration space
	$mathcal{C}(S)=V_{1}\times\dots\times V_{n}$ is partitioned
	\item an approximation $\mathscr{P}_{i}$ of the Pareto-optimal set.
	\item a set $test_{i}$ of at most $K$ configurations evaluated
	in the era. \comment{Nella era i o nella successiva? DP: non mi ricordo!}
	\end{itemize}
$K$ is a parameter that must be set before the algorithm runs.
\comment{Spiegare un po' l'impatto di K. DP: se c'e' tempo...} Note that $\mathcal{R}_{i}$ is a partition of the configuration space, i.e.
	\begin{align}\begin{array}{l}
		\bigcup_{R\in\mathcal{R}_i} R = \mathcal{C}(S) \mbox{ and}\\
		R' \cap R'' = \emptyset, \forall R',R'' \in \mathcal{R}_i
	\end{array}\end{align}

\comment{Prima di sbattere tutta la sfilza di steps, dovremmo dare
un'idea preliminare di cosa si andra' a fare. DP: se c'e' tempo bis}

@@

\subsection{Termination Condition}
The algorithm terminates at step $k$ if
\[
is_{TOT,k},is_{TOT,k-1},\dots,is_{TOT,k-\beta}\le\gamma\cdot is_{TOT,k-\left(\beta-1\right)}
\]
 where $\beta$ and $\gamma$ are selected by the experimenter (see
remark \ref{rem:termination}).
\begin{remark}
\label{pers02.rem:smaller_regions}The rule \ref{pers02.enu:K} of
the iteration rules says that the smaller is a region, the more ``crowded''
it is. It means that configurations in smaller regions are evaluated
with more ``attention'' (see figure \ref{pers02.fig:small_and_big}).

\begin{figure}[h]
\includegraphics[width=0.5\columnwidth]{img/small_and_big}

\caption{\label{pers02.fig:small_and_big}In this example 5 configurations
are evaluated in a small region and in a bigger region. In can be
seen that smaller region is more crowded.}

>>>>

\section{PS algorithm}
\secL{algorithm}
In this section we present PS, the algorithm that we propose to perform the Design Space Exploration (DSE). As any other DSE algorithm, its input is the parameterized system $S$ and a benchmark application $b$, while the output is an approximation of the Pareto-optimal set $\mathcal{P}\left(S,b\right)$.   
PS is iterative and we will describe, separately, the initialization phase, the operations performed inside each iteration and the condition that triggers the termination.

At a glance, at each iterations simulations are performed. Regions are classified based on the innovation introduced by their configurations in the Pareto-front so far calculated. The most interesting regions are split, so that each subregion will receive evaluation effort in the next iteration. On the contrary, uninteresting regions are merged.

@@

Each iteration is called \emph{era}. The $i$-th era is characterized by: 
	\begin{itemize}
	\item the set of regions $\mathcal{R}_{i}$ in which the configuration space
	$mathcal{C}(S)=V_{1}\times\dots\times V_{n}$ is partitioned
	\item an approximation $\mathscr{P}_{i}$ of the Pareto-optimal set.
	\item a set $test_{i}$ of at most $K$ configurations evaluated
	in the era.
	\end{itemize}
$K$ is a parameter that must be set before the algorithm runs.
\comment{Spiegare un po' l'impatto di K. DP: se c'e' tempo...} Note that $\mathcal{R}_{i}$ is a partition of the configuration space, i.e.
	\begin{align}\begin{array}{l}
		\bigcup_{R\in\mathcal{R}_i} R = \mathcal{C}(S) \mbox{ and}\\
		R' \cap R'' = \emptyset, \forall R',R'' \in \mathcal{R}_i
	\end{array}\end{align}

We now describe formally PS.

@@

\subsection{Termination Condition}
The algorithm terminates at step $k$ if
\[
is_{TOT,k},is_{TOT,k-1},\dots,is_{TOT,k-\beta}\le\gamma\cdot is_{TOT,k-\left(\beta-1\right)}
\]
 where $\beta$ and $\gamma$ are selected by the experimenter (see
remark \ref{rem:termination}).
\begin{remark}
\label{pers02.rem:smaller_regions}The rule \ref{pers02.enu:K} of
the iteration rules says that the smaller is a region, the more ``crowded''
it is. It means that configurations in smaller regions are evaluated
with more ``attention'' (see figure \ref{pers02.fig:small_and_big}).
\end{remark}

### 2014-07-27 concepts.tex
AA: aggiustato fino a initialization phase

\section{Underlying concepts and theoretical formulation}
\label{sec:statement_of_the_problem}

\comment{Theoretical suona troppo pomposo? DP: boh... AA: Esportare di
nuovo tutte le immagini libreoffice - svg - eps che viene meglio, come
si vede in \figR{3}. Aggiustare i nomi degli assi che in alcune figure
sono sbagliati. O forse basta esportarli col nuovo libreoffice che fa
meglio DP: hai ragione, ma forse siamo fuori tempo limite per queste
cose} 
In this section we provide the concepts and the formal
definitions that we will use later to express PS, the algorithm we
propose.  First, we give the definitions of Pareto set and Pareto
front that lie on the basis of Design Space Exploration is based on.
Then, we sketch the fundamental idea behind our algorithm. Finally, we
provide a formal definition of the regions, of their properties and of
the operations that will be performed on them by the algorithm.

@@

Let $S$ be a parameterized system with $n$ parameters. The generic
parameter $p_i, \ i \in \{1,2,\ldots,n\}$ can take any value in
the set $V_i$. A {\em configuration} $\mathbf{c}$ of the system
$S$ is a $n$-tuple $\langle v_1,v_2,\ldots,v_n \rangle$ in which
$v_i \in V_i$ is the value fixed for the parameter $p_i$. The {\em
configuration space} (or {\em design space}) of $S$ [which we will
indicate as $\mathcal{C}(S)$] is the complete range of possible
configurations [$\mathcal{C}(S) = V_1 \times V_2 \times \ldots
\times V_n$]. Naturally, not all the configurations of
$\mathcal{C}(S)$ can really be mapped on $S$, but only a subset 
is feasible. We call this subset {\em
feasible configuration space} $S$ and indicate it as
$\mathcal{C}^*(S)$].

Let $m$ be the number of objectives to be optimized (e.g. power,
cost, performance, etc.). An {\em evaluation function}
$E:\mathcal{C}^*(S)\times \mathcal{B} \longrightarrow \Re^m$ is a
function that associates each feasible configuration of $S$ with
an $m$-tuple of values corresponding to the objectives to be
optimized when any application belonging to the set of benchmarks
$\mathcal{B}$ is executed.

@@

\begin{definition}[Pareto Set]
\label{pers02.def:Pareto-set}
The {\em Pareto-optimal set} of $S$ for the application $b$ is the
set:
\[ \mathcal{P}(S,b) = \left\{ \mathbf{c} \in \mathcal{C}^*(S) \ | \nexists \ \mathbf{c}' \in \mathcal{C}^*(S), \mathbf{c}' \succ \mathbf{c} \right\} \]
\end{definition}
that is, the set of configurations $\mathbf{c} \in
\mathcal{C}^*(S)$ not dominated by any other configuration.
The configurations belonging to the Pareto-optimal set are called \emph{Pareto-optimal configurations}, while the {\em Pareto-optimal front} is their image, i.e. the set:
\[ \mathcal{P}_{F}(S,b) = \{ \mathbf{o} | \mathbf{o} = E(\mathbf{c},b), \mathbf{c} \in \mathcal{P}(S,b) \} \]

The aim of a Design Space Exploration (DSE) strategy is to give a
good approximation of the Pareto-optimal front for a system $S$ and an
application $b$, simulating as few configurations as possible.

@@

\comment{Mi sembra solo un cambio di notazione. DP: c'era un On the
contrary troppo aggressivio, ho cercato di alleggerire la cosa, ma dai
un'occhiata }

For each point of the Pareto front we can imagine the corresponding
configuration as a point of $P_{1}\times\dots\times P_{n}$
space. We will also refer to these points as ``parameter space representation'' of the
Pareto front. Figure~\ref{fig:psos} shows the relation between the objective space
(containing the Pareto front) and its parameter space representation.
In particular in this work we are interested in the relation between
the evolution of the points in the two spaces, e.g. how some
properties of the Pareto front points can reflect some features of the
corresponding points in the parameter space.

\begin{figure}[h]
\label{fig:psos}
\includegraphics[width=0.5\columnwidth]{img/Pareto_set_and_front}
\caption{Relation between Pareto set (its ``parameter space representation'')
and Pareto front (``objective space representation'').}
\end{figure}

@@

In the design process, there are often a lot of ``local decisions''
in which the designer can take advantage of well-known results or
heuristics. For example, if we consider a VLIW architecture in a configuration with very few registers, it is well-known that increasing the number of ALUs does not bring tangible better performances
but only cause an increase in area occupation. An intelligent exploration
algorithm would not waste time evaluating configurations with many
ALUs and few registers. If $p_{1}$ is the parameter ``number of
registers'' and $p_{2}$ is the parameter ``number of ALUs'', we
can say that the region:
\[
R=\left\{ \left.\left(v_{1},v_{2}\right)\in P_{1}\times P_{2}\right|v_{1}<s_{1},v_{2}>s_{2}\right\} 
\]
 (where $s_{1}$ and $s_{2}$ are some threshold values) is \emph{uninteresting}. This clearly shows that there are cases when, if a parameter lies in
a certain range, there are ranges of the other that are not worth exploring. Intuitively, the Cartesian product of these ranges is an ``uninteresting
 region''.

The problem is that, in complex scenarios, not all relations of this
kind are known in advance. Some dependency may exist but may be hidden
to the designer. Or they may be ``local'', i.e. two
parameters may be related by a dependency but only if their values
lie on some ranges (so we cannot treat them as they were related by
a ``complete'' dependency). 
Therefore, as stated in section \ref{sec:Related-work},
setting up an exploration trying to take into account as much dependencies
as possible is a hard task that is destined to be incompletely
accomplished. We need a methodology to ``automatically'' recognize
interesting or uninteresting regions without requiring a priori knowledge
to the designer, that is what our algorithm does.

We also consider, in measuring how much interesting a region is, the
\emph{innovation} that it adds. If, during a design space exploration,
a Pareto front is temporarily calculated, adding a new Pareto front
point near to the previous ones is not a considerable innovation,
because the way it fulfills the objectives is similar to the other
already evaluated configurations. On the contrary, adding a new Pareto
front point that is distant from previous ones is remarkable,
since it let the experimenter discover potential performance that have not yet considered before.

@@

\begin{definition}[Parameter Interval] Let $p_{i}$ be a parameter and $a_{i},b_{i}\in V_{i},a_{i}<b_{i}$.
A $p_{i}$- interval is
\[
\left[a_{i}..b_{i}\right]=\left[a_{i},b_{i}\right]\cap P_{i}
\]
i.e. taking the interval $\left[a_{i}..b_{i}\right]$
is the set of values of $V_{i}$ lying in $\left[a_{i},b_{i}\right]$)
\end{definition}

We consider only parameters $p_{i}$ with ordering, i.e. such
that $\forall a,b\in V_{i}$ it is possible
to say $a<b$ or $a=b$ or $a>b$.

%TODO: DESCRIVERE COME SI PENSA DI TRATTARE GLI SCENARI IN CUI ESISTONO
%PARAMETRI SENZA ORDINAMENTO{]}

@@

\begin{definition}[Contiguous intervals]
\label{pers02.def:Contiguous-intervals}
Two intervals are contiguous iff they do not overlap and merging them results in a new interval.More formally, let $\left[a_{i}..b_{i}\right]$ and $\left[x_{i}..y_{i}\right]$
be two $p_{i}$-intervals. They are said to be contiguous (see \figR{interval_and_region}(a) ) iff
\begin{align}\begin{array}{l}
	\left[a_{i}..b_{i}\right]\cap\left[x_{i}..y_{i}\right]=\emptyset \mbox{      and} \\
	\left[a_{i}..b_{i}\right]\cup\left[x_{i}..y_{i}\right]$ is a $p_{i}$-
interval or $\left[x_{i}..y_{i}\right]\cup\left[a_{i}..b_{i}\right]
\end{array}\end{align}
\end{definition}

@@

We now provide some rules to split and merge regions. They will be
useful when explaining our proposed exploration algorithm.
\begin{definition}[Splitting an interval]
\defL{splitting-a-region}
Given a $p_i$ interval $\left[a_{i}..b_{i}\right]$ with more than one element, i.e. $\left|\left[a_{i}..b_{i}\right]\right|>1$, the splitting operation produces as result two contiguous intervals $\left[a_{i}..c_{i}\right],\left[d_{i}..b_{i}\right]$
 such that
	\begin{align}
		\begin{cases}
		\left[a_{i}..c_{i}\right]\cap\left[d_{i}..b_{i}\right] & =\emptyset\\
		\left[a_{i}..c_{i}\right]\cup\left[d_{i}..b_{i}\right] & =\left[a_{i}..b_{i}\right]\\
		\left|\left[a_{i}..c_{i}\right]\right| & =\left\lceil \frac{\left|\left[a_{i}..b_{i}\right]\right|}{2}\right\rceil 
		\end{cases}
	\end{align}

@@

\begin{definition}[Contiguous regions]Let $R_{1}=\prod\left[a_{i}\dots b_{i}\right]$ and $R_{2}=\prod\left[x_{i}..y_{i}\right]$
be two regions. They are said to be contiguous iff they do not overlap and, merging them, a new region can be obtained. Formally, $R_1$ and $R_2$ are contiguous iff there exists a $j$ such that:
	\begin{align}\begin{array}{l}
		\left[a_{i}..b_{i}\right]=\left[x_{i}..y_{i}\right]$
	\mbox{ for } $i\neq j \mbox{      and} \\
		\left[a_{j}..b_{j}\right]$ and $\left[x_{j}..y_{j}\right] \mbox{ are contiguous intervals}
	\end{array}\end{align}
\noindent We call $p_j$ the \emph{contiguity parameter} for the two regions.
\end{definition}

Examples of contiguous and non contiguous regions are given in \figR{contiguous_regions}. 
\comment{Spiegare le figure e inserire a dx una figura con le contiguous regions mergiate. Le definizioni buttate cosi' rendono difficile la lettura a chi non sa ancora dove vogliamo andare a parare. Giustifica ogni definizione prima di darla. Ad es. quando parli di splittare, spiega che quando le regioni sono interessanti noi le splittiamo cosi' da esplorarle meglio. Al contrario quando le mergiamo. Mettere una sezione del tipo "Required operations at a glance" per giustificare tutte le definizioni prima di buttarle}.

	\begin{figure}[h]
		\includegraphics[width=0.5\columnwidth]{img/contiguous_regions}
		\figLC{contiguous_regions}{Contiguous and non contiguous
		regions.}
	\end{figure}

\begin{definition}[Merging intervals]
Let $\left[a_{i}\dots b_{i}\right]$ and $\left[x_{i}..y_{i}\right]$ be two contiguous $p_i$-intervals. The merge operation applied to them produces a new $p_i$-interval of the following form:
	\begin{align}
		\mu	 \left( \begin{array}{l}
				\left[a_{j}..y_{j}\right] \\
				\left[x_{j}..b_{j}\right]
		     \end{array} \right)
		=\begin{cases}
			\left[a_{j}..y_{j}\right] & \mbox{ if }b_{j}<x_{j}\\
			\left[x_{j}..b_{j}\right] & \mbox{ if }y_{j}<a_{j}
		\end{cases}
	\end{align}

>>>>

\section{Underlying concepts and theoretical formulation}
\secL{statement_of_the_problem}

In this section we provide the concepts and the formal
definitions that we will use later to express PS, the algorithm we
propose.  First, we give the definitions of Pareto set and Pareto
front that lie on the basis of Design Space Exploration.
Then, we sketch the fundamental idea behind our algorithm. Finally, we
provide a formal definition of the regions, of their properties and of
the operations that will be performed on them by the algorithm.

@@

Let $S$ be a parameterized system with $n$ parameters. The generic
parameter $p_i, \ i \in \{1,2,\ldots,n\}$ can take any value in
the set $V_i$. A {\em configuration} $\mathbf{c}$ of the system
$S$ is a $n$-tuple $\langle v_1,v_2,\ldots,v_n \rangle$ in which
$v_i \in V_i$ is the value fixed for the parameter $p_i$. The {\em
configuration space} (or {\em design space}) of $S$ [which we will
indicate as $\mathcal{C}(S)$] is the complete range of possible
configurations [$\mathcal{C}(S) = V_1 \times V_2 \times \ldots
\times V_n$]. Naturally, not all the configurations of
$\mathcal{C}(S)$ can really be mapped on $S$, but only a subset 
is feasible. We call this subset {\em
feasible configuration space} of $S$ and indicate it as
$\mathcal{C}^*(S)$].

Let $m$ be the number of objectives to be optimized (e.g. power,
cost, performance, etc.). An {\em evaluation function}
$E:\mathcal{C}^*(S)\times \mathcal{B} \longrightarrow \Re^m$ associates each feasible configuration of $S$ to
an $m$-tuple of values corresponding to the objectives when any application belonging to the set of benchmarks
$\mathcal{B}$ is executed.

@@

\begin{definition}[Pareto set and Pareto front]
\defL{pareto-set}
The {\em Pareto-optimal set} of $S$ (or, shortly, Pareto-set) for the application $b$ is the
set:
\[ \mathcal{P}(S,b) = \left\{ \mathbf{c} \in \mathcal{C}^*(S) \ | \nexists \ \mathbf{c}' \in \mathcal{C}^*(S), \mathbf{c}' \succ \mathbf{c} \right\} \]
\end{definition}
that is, the set of configurations $\mathbf{c} \in
\mathcal{C}^*(S)$ not dominated by any other configuration.
The configurations belonging to the Pareto-optimal set are called \emph{Pareto-optimal configurations}, while the {\em Pareto-optimal front} (or, shortly, Pareto-front) is their image, i.e. the set:
\[ \mathcal{P}_{F}(S,b) = \{ \mathbf{o} | \mathbf{o} = E(\mathbf{c},b), \mathbf{c} \in \mathcal{P}(S,b) \} \]

The aim of a Design Space Exploration (DSE) strategy is to give a
good approximation of the Pareto-optimal set for a system $S$ and an
application $b$, simulating as few configurations as possible.

@@

As we will show later, at each iteration, PS calculates an approssimation $\mathscr{P}_i$ of the Pareto-set that approaches it as iterations go on. The image of $\mathscr{P}_i$ is an approssimation of the Pareto-front.
The simulation effort is distributed more on the regions that provide configurations in $\mathscr{P}_i$ which introduce more \emph{innovation}, i.e. whose image in the objective space is more distant, thus ``more different'', from the previous Pareto-front approssimation point.

% A PROPOSITO DI QUESTO DUBBIO:
%\comment{AA: Mi sembra solo un cambio di notazione. DP: c'era un On the contrary troppo aggressivio, ho cercato di alleggerire la cosa, ma dai un'occhiata }
% HO CANCELLATO TUTTA QUESTA PARTE.
% For each point of the Pareto front we can imagine the corresponding configuration as a point of $P_{1}\times\dots\times P_{n}$ space. We will also refer to these points as ``parameter space representation'' of the Pareto front. Figure~\ref{fig:psos} shows the relation between the objective space (containing the Pareto front) and its parameter space representation. In particular in this work we are interested in the relation between the evolution of the points in the two spaces, e.g. how some properties of the Pareto front points can reflect some features of the corresponding points in the parameter space.

\begin{figure}[t]
\label{fig:psos}
\includegraphics[width=0.5\columnwidth]{img/Pareto_set_and_front}
\caption{Relation between Pareto-set and Pareto front.}
\end{figure}

@@

In the design process, there are often a lot of ``local decisions''
in which the designer can take advantage of well-known results or
heuristics. For example, if we consider a VLIW architecture in a configuration with very few registers, it is well-known that increasing the number of ALUs does not bring tangible better performances
but only cause an increase in area occupation. An intelligent exploration
algorithm would not waste time evaluating configurations with many
ALUs and few registers. If $p_{1}$ is the parameter ``number of
registers'' and $p_{2}$ is the parameter ``number of ALUs'', we
can say that the region:
\[
R=\left\{ \left.\left(v_{1},v_{2}\right)\in P_{1}\times P_{2}\right|v_{1}<s_{1},v_{2}>s_{2}\right\} 
\]
 (where $s_{1}$ and $s_{2}$ are some threshold values) is \emph{uninteresting}. This clearly shows that there are cases when, if a parameter lies in
a certain range, there are ranges of the other one that are not worth being explored. Intuitively, the cartesian product of these ranges is an \emph{uninteresting region}.

The problem is that, in complex scenarios, not all the relations of this
kind are known in advance. Some dependency may be hidden
to the designer.
Therefore, as stated in section \ref{sec:Related-work},
setting up an exploration trying to take into account as much dependencies
as possible is a hard task that is destined to be incompletely
accomplished. We need a methodology to ``automatically'' recognize
interesting or uninteresting regions without requiring a priori knowledge
to the designer, that is what our algorithm does.

We also consider, in measuring how much interesting a region is, the
\emph{innovation} that it adds. If, during a design space exploration,
a Pareto front approximation is temporarily calculated, adding a new Pareto front
point near to the previous ones is not a considerable innovation,
because the way it fulfills the objectives is similar to the other
already evaluated configurations. On the contrary, adding a new Pareto
front point that is distant from previous ones is remarkable,
since it let the experimenter discover potential performance that have not yet considered before.
At each iteration, PS ``zooms in'' the most interesting regions while ``zooming out'' from the rest of configuration space. In order to implement this behavior algorithmically, at each iteration PS assigns the same number $N$ of simulations to each region. If a region $R$ is interesting, it is split in $k$ subregions before the next iteration starts. Therefore, in the next iteration, $N$ simulations are assigned to each subregion. As a result, we are populating the original region $R$ with $k\cdot N$ simulations. On the contrary, uninteresting regions are merged, so that each one will receive less effort in the next iteration. In other words, ares of the configuration space that are more promising are split in small subregions, while the other ones are blended in big regions. As illustrated in \figR{small_and_big}, assigning $N$ simulations to each region makes small ones well populated while bigger ones sparsely explored.

	\begin{figure}[t]
	\includegraphics[width=0.5\columnwidth]{img/small_and_big}
	\figLC{small_and_big}{In this example 5 configurations
	are evaluated in a small region and in a bigger region. In can be
	seen that smaller region is more crowded.}
	\end{figure}

@@

Since regions will be defined as cartesian products of parameter intervals, we need the following definition.
\begin{definition}[Parameter interval] Let $p_{i}$ be a parameter and $a_{i},b_{i}\in V_{i},a_{i}<b_{i}$.
A $p_{i}$- interval is
	\[
	\left[a_{i}..b_{i}\right]=\left[a_{i},b_{i}\right]\cap P_{i}
	\]
i.e. taking the interval $\left[a_{i}..b_{i}\right]$
is the set of values of $V_{i}$ lying in $\left[a_{i},b_{i}\right]$)
\end{definition}

@@

We consider only parameters $p_{i}$ with ordering, i.e. such
that $\forall a,b\in V_{i}$ it is possible
to say $a<b$ or $a=b$ or $a>b$. 
The following definition of interval contiguity will be the base of the concept of region contiguity, used to define merging and splitting operations.
\begin{definition}[Contiguous intervals]
\label{pers02.def:Contiguous-intervals}
Two intervals are contiguous iff they do not overlap and merging them results in a new interval.More formally, let $\left[a_{i}..b_{i}\right]$ and $\left[x_{i}..y_{i}\right]$
be two $p_{i}$-intervals. They are said to be contiguous (see \figR{interval_and_region}(a) ) iff
\begin{align}\begin{array}{l}
	\left[a_{i}..b_{i}\right]\cap\left[x_{i}..y_{i}\right]=\emptyset \mbox{      and} \\
	\left[a_{i}..b_{i}\right]\cup\left[x_{i}..y_{i}\right]$ is a $p_{i}$-
interval or $\left[x_{i}..y_{i}\right]\cup\left[a_{i}..b_{i}\right]
\end{array}\end{align}
\end{definition}

@@

As already explain, we will split interesting regions and merge uninteresting ones. The region splitting and merging operations are defined starting from the corresponding operations defined on intervals.
\begin{definition}[Splitting an interval]
\defL{splitting-a-region}
Given a $p_i$ interval $\left[a_{i}..b_{i}\right]$ with more than one element, i.e. $\left|\left[a_{i}..b_{i}\right]\right|>1$, the splitting operation produces as result two contiguous intervals $\left[a_{i}..c_{i}\right],\left[d_{i}..b_{i}\right]$
 such that
	\begin{align}
		\begin{cases}
		\left[a_{i}..c_{i}\right]\cap\left[d_{i}..b_{i}\right] & =\emptyset\\
		\left[a_{i}..c_{i}\right]\cup\left[d_{i}..b_{i}\right] & =\left[a_{i}..b_{i}\right]\\
		\left|\left[a_{i}..c_{i}\right]\right| & =\left\lceil \frac{\left|\left[a_{i}..b_{i}\right]\right|}{2}\right\rceil 
		\end{cases}
	\end{align}

@@

The merging operation will be defined only on region pairs that satisfy the conditions provided in the following definition.
\begin{definition}[Contiguous regions]Let $R_{1}=\prod\left[a_{i}\dots b_{i}\right]$ and $R_{2}=\prod\left[x_{i}..y_{i}\right]$
be two regions. They are said to be contiguous iff they do not overlap and, merging them, a new region can be obtained. Formally, $R_1$ and $R_2$ are contiguous iff there exists a $j$ such that:
	\begin{align}\begin{array}{l}
		\left[a_{i}..b_{i}\right]=\left[x_{i}..y_{i}\right]$
	\mbox{ for } $i\neq j \mbox{      and} \\
		\left[a_{j}..b_{j}\right]$ and $\left[x_{j}..y_{j}\right] \mbox{ are contiguous intervals}
	\end{array}\end{align}
\noindent We call $p_j$ the \emph{contiguity parameter} for the two regions.
\end{definition}

Examples of contiguous and non contiguous regions are given in \figR{contiguous_regions}. The two regions at the top are contiguous because their union can be expressed as the cartesian product of parameter intervals, thus they can be merged to obtain a new bigger region. The same applies also to the regions on the right. On the contrary, the two regions on the left (or of the two regions at the bottom) cannot be merged.

	\begin{figure}[h]
		\includegraphics[width=0.5\columnwidth]{img/contiguous_regions}
		\figLC{contiguous_regions}{Contiguous and non contiguous regions. \comment{Inserire a dx una figura con le contiguous regions mergiate.}	}
	\end{figure}

We now describe formally the merging operations.
\begin{definition}[Merging intervals]
Let $\left[a_{i}\dots b_{i}\right]$ and $\left[x_{i}..y_{i}\right]$ be two contiguous $p_i$-intervals. The merge operation applied to them produces a new $p_i$-interval of the following form:
	\begin{align}
		\mu	 \left( \begin{array}{l}
				\left[a_{j}..y_{j}\right] \\
				\left[x_{j}..b_{j}\right]
		     \end{array} \right)
		=\begin{cases}
			\left[a_{j}..y_{j}\right] & \mbox{ if }b_{j}<x_{j}\\
			\left[x_{j}..b_{j}\right] & \mbox{ if }y_{j}<a_{j}
		\end{cases}
	\end{align}

### 2014-07-27 introduction.tex
AA: aggiustato fino a initialization phase

The design of an embedded system requires different conflicting
objectives (energy consumption, performance, area) to be fulfilled.
Several factors are involved in this goal and real cases show how
often no analytical results are known to predict the relation between
system configurations and objectives. In such cases a high level
system simulation is often the only way to have a picture of how system
parameters impact on the objectives.  A key problem is that the size
of the design space to be simulated grows with the product of
cardinalities of each parameter, easily resulting in an intractable number
of simulations. In these cases a ``Design Space
Exploration'' strategy is required, i.e. a methodology to evaluate
only a small subset of all possible configurations while maintaining a
good level of efficiency and accuracy~\cite{surviving_soc}.

The final result of a design space exploration is a subset of the
configurations called Pareto Set (\cite{pareto}) (See also Definition
\ref{pers02.def:Pareto-set}). It represents the trade-offs between
objectives, so that for each configuration of the Pareto Set there is no
other configuration performing better for all the objectives
considered. Once these few promising configurations have been obtained, a
subsequent step of the design flow can eventually afford a more
accurate low level simulation.

In this work we present a multiobjective design exploration strategy
that introduces the concept of Parameter Space Representation of
Pareto Set (PS). The main idea is focusing on the ``innovation''
introduced in the Pareto sets and then accordingly distributing the
exploration effort in regions created in the multidimensional space of
system parameters. The aim of the PS approach is to propose a new
perspective on the configuration space, where parameter
interdependencies play the fundamental role of making some regions
more capable of generating hardly predictable solutions: catching this
sort of ``chaoticity'' in the parameter space is the main idea behind
the strategy that is being introduced in this work.

\section{Related work and Contribution}
\label{sec:Related-work}
Many different design space exploration algorithms have been proposed
in literature.  The motivations of an exploration
algorithm are rather heuristic, have some form of arbitrariness and,
to a large extend, intuition lies behind them.

@@

Many studies, as \cite{coello_easmop} and others, focus on genetic approaches to solve multiobjective
optimization problems. Genetic approaches
have many advantages: they provide good accuracy, they are customizable and very general
and require no a-priori knowledge to the designer.  The strong
point of genetic approaches can be summarized saying that they consist
in exploration that is sufficiently broad (the mutation avoids being rigidly restricted to limited parts of parameter space), and, at the same time, not too scattered as it is guided by the performances of
the already evaluated configurations. We claim that the
approach presented in this paper has most of the benefits of the genetic
approaches, although the rationale is completely different.

The paper is organized as follows: in
Section~\ref{sec:statement_of_the_problem} we formally introduce the
main concepts and the definitions required to apply the proposed strategy. Next in
Section~\ref{sec:algorithm} we present the PS algorithm in detail.
Finally in Section~\ref{sec:ee} we show a case study involving the
exploration of the hardware/software parameters of a Very Long
 Instruction Word (VLIW) architecture, performing a qualitative
 and quantitative comparison of
PS agains the state-of-art of multiobjective genetic based approaches.

>>>>

The design of an embedded system requires different conflicting
objectives (energy consumption, performance, area) to be fulfilled.
Several factors are involved in this goal and real cases show how
often no analytical results are known to predict the relation between
system configurations and objectives. In such cases a high level
system simulation is the only way to have a picture of how system
parameters impact on the objectives.  A key problem is that the size
of the design space to be simulated grows with the product of
cardinalities of each parameter, resulting in an intractable number
of simulations. In these cases a Design Space
Exploration (DSE) strategy is required, i.e. an algorithm that selects
only a tractable subset of all possible configurations to simulate while providing, at the end, configurations that are close to the optimum and that are worth evaluating, in a following step of the design process, in a more thorough way.

%The final result of a design space exploration is a subset of the configurations called \emph{Pareto set} (\cite{pareto}, see also \defR{pareto-set}). The corresponding values of the objectives for each configuration of the Pareto Set is called \emph{Pareto front}: it represents the trade-offs between objectives, i.e.  for each point of the Pareto front there is no other configuration performing better for all the objectives considered. Once these few promising configurations have been obtained, a subsequent step of the design flow can eventually afford a more accurate low level simulation.

In this work we present a multiobjective design exploration algorithm
that introduces the concept of interesting and uninteresting regions of the configuration space.
\comment{AA: prima era ``algorithm
that introduces the concept of Parameter Space Representation of
Pareto front''. Se io fossi un reviewer, direi: guardi, questo io non credo, anche pecche' non siamo noi che introduciamo il concetto di parameter space representation del pareto set. Il concetto esiste gia' dal 1896 e si chiama ``Pareto-set''! Ho cambiato il claim dicendo che introduciamo interesting/uninteresting regions, che mi sembra piu' rispondente alla verita'. Cosi' ti va bene?} 
The algorithm is iterative. The main idea can be simplified as follows. At each iteration, the configuration space is divided in regions. A score is given to each region based on the \emph{innovation} provided by its configurations in the Pareto front w.r.t. the previous iterations. The regions with high score are considered the most \emph{interesting} and more simulations are distributed there rather than in the rest of the configuration space. Since the fundamental operations consist in the formation of regions in the Parameter Space (aka configuration space), we call our algorithm PS.

\comment{AA: Le seguenti referenze a cose di psicologia cognitiva potrebbero essere interpretate come una figata o una cagata. Tu come le interpreti?}
Our algorithm mimics the attention shift in human beings. Authors in \cite{attention} models human attention as a single resource with a limited budget that must be distributed among tasks, preferring the task which the focus is on.
 Authors in \cite{spatial_attention} describes the visual attention as a spotlight that focuses, each time, in a particular region of the visual space in order to process it more deeply. Similarly, PS focuses its ``spotlight'' only to the interesting regions and explore them more thoroughly distributing the simulation budget on them.
% Ispirzione: http://en.wikipedia.org/wiki/Attentional_shift

With PS, we propose a new
perspective on the configuration space, where parameter
interdependencies play the fundamental role of making some regions
more capable of generating hardly predictable solutions: catching this
sort of ``chaoticity'' in the parameter space is the main idea behind
the strategy that is being introduced in this work.

The paper is organized as follows. First, \secR{Related-work} places our work in the context of related effort.
Then we give a theoretical formulation of the algorithm. In particular, in \secR{statement_of_the_problem} we provide formal definitions of the concepts and the operations that will be used to describe PS algorithm. In \secR{algorithm} we give the theoretical description of PS. 
Finally in Section~\ref{sec:ee} we show a case
study involving the exploration of the hardware/software parameters of
a Very Long Instruction Word (VLIW) architecture, performing a
qualitative and quantitative comparison of PS against the state-of-art
of multiobjective genetic based approaches.

\section{Related work}
\secL{Related-work}
Many different design space exploration algorithms have been proposed
in literature.  The motivations of an exploration
algorithm are rather heuristic, have some form of arbitrariness and,
to a large extend, intuition lies behind them.

@@

Many studies, as \cite{coello_easmop} and others, focus on genetic approaches to solve multiobjective
optimization problems. Genetic approaches
have many advantages: they provide good accuracy, they are customizable and very general
and require no a-priori knowledge to the designer.  The strong
point of genetic approaches can be summarized saying that they consist
in exploration that is sufficiently broad (the mutation avoids being rigidly restricted to limited parts of parameter space), and, at the same time, not too scattered as it is guided by the performances of
the already evaluated configurations. In this work we will compare 
the proposed approach against a well known Multi-Objective Genetic
Algoritm~\cite{knowles_techrep06}.
We claim that the approach presented in this paper has most of the benefits of the genetic
approaches, although the rationale is completely different.
